Complementary Filters Shaping Using \(\mathcal{H}_\infty\) Synthesis

### 2019-09-26 paper/paper.tex
One figure was missing the Frequency axis
Review - Change test about robustnesss

% Created 2019-08-20 mar. 16:57
% Intended LaTeX compiler: pdflatex
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{siunitx}
\usepackage{amsmath,amssymb,amsfonts, cases}
\usepackage[noadjust,space,compress]{cite}
\usepackage{tabularx,siunitx,booktabs}
\usepackage{algorithmic, graphicx, textcomp}
\usepackage{xcolor, import, hyperref}
\usepackage[USenglish, english]{babel}
\setcounter{footnote}{1}
\renewcommand{\citedash}{--}
\IEEEoverridecommandlockouts
\author{Dehaeze Thomas\textsuperscript{1,2,$\dagger$}, Verma Mohit\textsuperscript{1,3} and Collette Christophe\textsuperscript{1,3}  \thanks{\textsuperscript{1} Precision Mechatronics Laboratory, A\&M Department, University of Liege, Belgium} \thanks{\textsuperscript{2} European Synchrotron Radiation Facility, Grenoble, France} \thanks{\textsuperscript{3} BEAMS Department, Free University of Brussels, Belgium} \thanks{\textsuperscript{$\dagger$} Corresponding Author. Email: {\tt\small thomas.dehaeze@esrf.fr}}}
\date{2019-08-20}
\title{Complementary Filters Shaping Using \(\mathcal{H}_\infty\) Synthesis}
\hypersetup{
 pdfauthor={Dehaeze Thomas\textsuperscript{1,2,$\dagger$}, Verma Mohit\textsuperscript{1,3} and Collette Christophe\textsuperscript{1,3}  \thanks{\textsuperscript{1} Precision Mechatronics Laboratory, A\&M Department, University of Liege, Belgium} \thanks{\textsuperscript{2} European Synchrotron Radiation Facility, Grenoble, France} \thanks{\textsuperscript{3} BEAMS Department, Free University of Brussels, Belgium} \thanks{\textsuperscript{$\dagger$} Corresponding Author. Email: {\tt\small thomas.dehaeze@esrf.fr}}},
 pdftitle={Complementary Filters Shaping Using \(\mathcal{H}_\infty\) Synthesis},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.2 (Org mode 9.2.5)},
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgf075900}
\label{sec:introduction}
A set of filters is said to be complementary if the sum of their transfer functions is equal to one at all frequencies.
These filters are used when two or more sensors are measuring the same physical quantity with different noise characteristics. Unreliable frequencies of each sensor are filtered out by the complementary filters and then combined to form a super sensor giving a better estimate of the physical quantity over a wider bandwidth.
This technique is called sensor fusion and is used in many applications.\par
In \cite{zimmermann92_high_bandw_orien_measur_contr,corke04_inert_visual_sensin_system_small_auton_helic,min15_compl_filter_desig_angle_estim}, various sensors (accelerometers, gyroscopes, vision sensors, etc.) are merged using complementary filters for the attitude estimation of Unmanned Aerial Vehicles (UAV).
In \cite{collette15_sensor_fusion_method_high_perfor}, several sensor fusion configurations using different types of sensors are discussed in order to increase the control bandwidth of active vibration isolation systems.
Furthermore, sensor fusion is used in the isolation systems of the Laser Interferometer Gravitational-Wave Observator (LIGO) to merge inertial sensors with relative sensors
\cite{matichard15_seism_isolat_advan_ligo,hua04_polyp_fir_compl_filter_contr_system}. \par
As the super sensor noise characteristics largely depend on the complementary filter norms, their proper design is of primary importance for sensor fusion.
In \cite{corke04_inert_visual_sensin_system_small_auton_helic,jensen13_basic_uas,min15_compl_filter_desig_angle_estim}, first and second order analytical formulas of complementary filters have been presented.
Higher order complementary filters have been used in
\cite{shaw90_bandw_enhan_posit_measur_using_measur_accel,zimmermann92_high_bandw_orien_measur_contr,collette15_sensor_fusion_method_high_perfor}.
In \cite{jensen13_basic_uas}, the sensitivity and complementary sensitivity transfer functions of a feedback architecture have been proposed to be used as complementary filters. The design of such filters can then benefit from the classical control theory developments.
Linear Matrix Inequalities (LMIs) are used in \cite{pascoal99_navig_system_desig_using_time} for the synthesis of complementary filters satisfying some frequency-like performance.
Finally, a synthesis method of high order Finite Impulse Response (FIR) complementary filters using convex optimization has been developed in \cite{hua05_low_ligo,hua04_polyp_fir_compl_filter_contr_system}. \par
Although many design methods of complementary filters have been proposed in the literature, no simple method that allows to shape the norm of the complementary filters is available.
This paper presents a new design method of complementary filters based on \(\mathcal{H}_\infty\) synthesis.
This design method permits to easily shape the norms of the generated filters.\par
The section \ref{sec:requirements} gives a brief overview of sensor fusion using complementary filters and explains how the typical requirements for such fusion can be expressed as upper bounds on the filters norms.
In section \ref{sec:hinf_method}, a new design method for the shaping of complementary filters using \(\mathcal{H}_\infty\) synthesis is proposed.
In section \ref{sec:application_ligo}, the method is used to design complex complementary filters that are used for sensor fusion at the LIGO.
Our conclusions are drawn in the final section.

\section{Complementary Filters Requirements}
\label{sec:orgcfd59fa}
\label{sec:requirements}
\subsection{Sensor Fusion Architecture}
\label{sec:orgc84edaa}
\label{sec:sensor_fusion}

Let's consider two sensors measuring the same physical quantity \(x\) with dynamics \(G_1(s)\) and \(G_2(s)\), and with noise characteristics \(n_1\) and \(n_2\).

@@

\subsection{Noise Sensor Filtering}
\label{sec:org3904fad}
\label{sec:noise_filtering}

@@

The estimation error \(\delta x\) is defined by \eqref{eq:estimate_error}.
\begin{equation}
\label{eq:estimate_error}
  \delta x \triangleq \hat{x} - x = H_1 n_1 + H_2 n_2
\end{equation}

@@

\subsection{Robustness of the Fusion}
\label{sec:org1955823}
\label{sec:fusion_robustness}

In practical systems, the sensor dynamics always have some level of uncertainty and cannot be inverted perfectly such that \(G_i(s) = 1\).

This can be represented as an input multiplicative uncertainty as shown in Fig. \ref{fig:sensor_fusion_dynamic_uncertainty} where \(\Delta_i\) is any stable transfer function satisfying \(|\Delta_i(j\omega)| \le 1,\ \forall\omega\), and \(|w_i(s)|\) is a weight representing the magnitude of the uncertainty.

@@

The super sensor dynamics \eqref{eq:super_sensor_dyn_uncertainty} is no longer equal to \(1\) and is now dependent on the sensor dynamics uncertainty weights \(w_i(s)\) as well as on the complementary filters \(H_i(s)\).
\begin{equation}
\label{eq:super_sensor_dyn_uncertainty}
  \frac{\hat{x}}{x} = 1 + w_1(s) H_1(s) \Delta_1(s) + w_2(s) H_2(s) \Delta_2(s)
\end{equation}

In order to limit the phase and gain uncertainty of the super sensor, one may want to design the complementary filters such that \eqref{eq:max_uncertainty_super_sensor} is satisfied.
\begin{equation}
\label{eq:max_uncertainty_super_sensor}
  \begin{aligned}
    & \left|w_1 H_1 \Delta_1\right| + \left|w_2 H_2 \Delta_2\right| \le \epsilon \quad \forall\omega,\ \forall \Delta_1, \forall \Delta_2\\
    \Leftrightarrow & \left|w_1 H_1\right| + \left|w_2 H_2\right| \le \epsilon \quad \forall\omega
  \end{aligned}
\end{equation}

Condition \eqref{eq:max_uncertainty_super_sensor} is equivalent to bounding the uncertainty set of the super sensor dynamics in the complex plane by a circle centered at \(1\) with a radius equal to \(\epsilon\) (Fig. \ref{fig:uncertainty_set_super_sensor}).

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/uncertainty_set_super_sensor.pdf}
\caption{\label{fig:uncertainty_set_super_sensor}
Uncertainty set of the super sensor dynamics}
\end{figure}

The maximum phase added by the super sensor uncertainty \(\Delta\phi\) is then
\begin{equation}
\label{eq:max_phase_uncertainty}
    \Delta \phi = \arcsin\left( \epsilon \right)
\end{equation}

Limiting the phase added by the super sensor to \(\Delta \phi = \SI{30}{\degree}\) requires that \(H_1(s)\) and \(H_2(s)\) are designed such that \eqref{eq:max_uncertainty_super_sensor} is satisfied with \(\epsilon = \sin(\Delta\phi = \SI{30}{\degree}) = 0.5\).
Thus the norm of the complementary filter \(|H_i|\) should be made small at frequencies where the dynamic uncertainty \(|w_i|\) of sensor \(i\) is large.

\section{Complementary Filters Shaping using \(\mathcal{H}_\infty\) Synthesis}
\label{sec:org8eb875c}
\label{sec:hinf_method}
As shown in Sec. \ref{sec:requirements}, the performance and robustness of the sensor fusion architecture depends on the complementary filters norms.
Therefore, the development of a synthesis method of complementary filters that allows the shaping of their norm is necessary.
\subsection{Shaping of Complementary Filters using \(\mathcal{H}_\infty\) synthesis}
\label{sec:orga1b497a}
\label{sec:hinf_synthesis}
The synthesis objective is to shape the norm of two filters \(H_1(s)\) and \(H_2(s)\) while ensuring their complementary property \eqref{eq:comp_filter}.
This is equivalent as to finding stable transfer functions \(H_1(s)\) and \(H_2(s)\) such that conditions \eqref{eq:comp_filter_problem_form} are satisfied.
\begin{subequations}
\label{eq:comp_filter_problem_form}
  \begin{align}
  & H_1(s) + H_2(s) = 1 \label{eq:hinf_cond_complementarity} \\
  & |H_1(j\omega)| \le \frac{1}{|W_1(j\omega)|} \quad \forall\omega \label{eq:hinf_cond_h1} \\
  & |H_2(j\omega)| \le \frac{1}{|W_2(j\omega)|} \quad \forall\omega \label{eq:hinf_cond_h2}
  \end{align}
\end{subequations}
where \(W_1(s)\) and \(W_2(s)\) are two weighting transfer functions that are chosen to shape the norms of the corresponding filters.

@@

\subsection{Weighting Functions Design}
\label{sec:orgdf5910c}
\label{sec:hinf_weighting_func}
The proper design of the weighting functions is of primary importance for the success of the presented complementary filters \(\mathcal{H}_\infty\) synthesis.

@@

\subsection{Validation of the proposed synthesis method}
\label{sec:org32ec287}
\label{sec:hinf_example}
Let's validate the proposed design method of complementary filters with a simple example where two complementary filters \(H_1(s)\) and \(H_2(s)\) have to be designed such that:
\begin{itemize}
\item the merging frequency is around \(\SI{10}{Hz}\)
\item the slope of \(|H_1(j\omega)|\) is \(-2\) above \(\SI{10}{Hz}\)
\item the slope of \(|H_2(j\omega)|\) is \(+3\) below \(\SI{10}{Hz}\)
\item the gain of both filters is equal to \(10^{-3}\) away from the merging frequency
\end{itemize}

@@

\subsection{Synthesis of Three Complementary Filters}
\label{sec:orgdfc880e}
\label{sec:hinf_three_comp_filters}
Some applications may require to merge more than two sensors.
In such a case, it is necessary to design as many complementary filters as the number of sensors used.
The synthesis problem is then to compute \(n\) stable transfer functions \(H_i(s)\) such that \eqref{eq:hinf_problem_gen} is satisfied.
\begin{subequations}
\label{eq:hinf_problem_gen}
  \begin{align}
  & \sum_{i=0}^n H_i(s) = 1 \label{eq:hinf_cond_compl_gen} \\
  & \left| H_i(j\omega) \right| < \frac{1}{\left| W_i(j\omega) \right|}, \quad \forall \omega,\ i = 1 \dots n \label{eq:hinf_cond_perf_gen}
  \end{align}
\end{subequations}
The synthesis method is generalized here for the synthesis of three complementary filters using the architecture shown in Fig. \ref{fig:comp_filter_three_hinf}.

@@

\section{Application: Design of Complementary Filters used in the Active Vibration Isolation System at the LIGO}
\label{sec:org58b0928}
\label{sec:application_ligo}
Several complementary filters are used in the active isolation system at the LIGO \cite{hua05_low_ligo,hua04_polyp_fir_compl_filter_contr_system}.
The requirements on those filters are very tight and thus their design is complex.
The approach used in \cite{hua05_low_ligo} for their design is to write the synthesis of complementary FIR filters as a convex optimization problem.
The obtained FIR filters are compliant with the requirements. However they are of very high order so their implementation is quite complex.

The effectiveness of the proposed method is demonstrated by designing complementary filters with the same requirements as the one described in \cite{hua05_low_ligo}.
\subsection{Complementary Filters Specifications}
\label{sec:org0bf5de3}
\label{sec:ligo_specifications}
The specifications for one pair of complementary filters used at the LIGO are summarized below (for further details, refer to \cite{hua04_polyp_fir_compl_filter_contr_system}):
\begin{itemize}
\item From \(0\) to \(\SI{0.008}{Hz}\), the magnitude of the filter's transfer function should be less or equal to \(8 \times 10^{-4}\)
\item Between \(\SI{0.008}{Hz}\) to \(\SI{0.04}{Hz}\), the filter should attenuate the input signal proportional to frequency cubed
\item Between \(\SI{0.04}{Hz}\) to \(\SI{0.1}{Hz}\), the magnitude of the transfer function should be less than \(3\)
\item Above \(\SI{0.1}{Hz}\), the magnitude of the complementary filter should be less than \(0.045\)
\end{itemize}

The specifications are represented by the dashed lines in Fig. \ref{fig:ligo_weights}.

\subsection{Weighting Functions Design}
\label{sec:org0a19352}
\label{sec:ligo_weights}
The weighting functions should be designed such that their inverse magnitude is as close as possible to the specifications in order to not over-constrain the synthesis problem.
However, the order of each weight should stay reasonably small in order to reduce the computational costs of the optimization problem as well as for the physical implementation of the filters.

A Type I Chebyshev filter of order \(20\) is used as the weighting transfer function \(w_L(s)\) corresponding to the low pass filter. For the one corresponding to the high pass filter \(w_H(s)\), a \(7^{\text{th}}\) order transfer function is designed.
The magnitudes of the weighting functions inverse are shown in Fig. \ref{fig:ligo_weights}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/ligo_weights.pdf}
\caption{\label{fig:ligo_weights}
Specifications and weighting functions magnitude used for \(\mathcal{H}_\infty\) synthesis}
\end{figure}

\subsection{\(\mathcal{H}_\infty\) Synthesis}
\label{sec:orgf481d28}
\label{sec:ligo_results}
\(\mathcal{H}_\infty\) synthesis is performed using the architecture shown in Fig. \ref{eq:generalized_plant}.
The complementary filters obtained are of order \(27\).
In Fig. \ref{fig:comp_fir_ligo_hinf}, their bode plot is compared with the FIR filters of order 512 obtained in \cite{hua05_low_ligo}.
They are found to be very close to each other and this shows the effectiveness of the proposed synthesis method.

@@

\section{Conclusion}
\label{sec:orgc26b54f}
  \label{sec:conclusion}
This paper has shown how complementary filters can be used to combine multiple sensors in order to obtain a super sensor.
Typical specification on the super sensor noise and on the robustness of the sensor fusion has been shown to be linked to the norm of the complementary filters.
Therefore, a synthesis method that permits the shaping of the complementary filters norms has been proposed and has been successfully applied for the design of complex filters.
Future work will aim at further developing this synthesis method for the robust and optimal synthesis of complementary filters used in sensor fusion.

\section{Acknowledgment}
\label{sec:org698db58}
This research benefited from a FRIA grant from the French Community of Belgium.

>>>>

% Created 2019-09-26 jeu. 18:26
% Intended LaTeX compiler: pdflatex
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{siunitx}
\usepackage{amsmath,amssymb,amsfonts, cases}
\usepackage[noadjust,space,compress]{cite}
\usepackage{tabularx,siunitx,booktabs}
\usepackage{algorithmic, graphicx, textcomp}
\usepackage{xcolor, import, hyperref}
\usepackage[USenglish, english]{babel}
\setcounter{footnote}{1}
\renewcommand{\citedash}{--}
\IEEEoverridecommandlockouts
\author{Dehaeze Thomas\textsuperscript{1,2,$\dagger$}, Verma Mohit\textsuperscript{1,3} and Collette Christophe\textsuperscript{1,3}  \thanks{\textsuperscript{1} Precision Mechatronics Laboratory, A\&M Department, University of Liege, Belgium} \thanks{\textsuperscript{2} European Synchrotron Radiation Facility, Grenoble, France} \thanks{\textsuperscript{3} BEAMS Department, Free University of Brussels, Belgium} \thanks{\textsuperscript{$\dagger$} Corresponding Author. Email: {\tt\small thomas.dehaeze@esrf.fr}}}
\date{2019-09-26}
\title{Complementary Filters Shaping Using \(\mathcal{H}_\infty\) Synthesis}
\hypersetup{
 pdfauthor={Dehaeze Thomas\textsuperscript{1,2,$\dagger$}, Verma Mohit\textsuperscript{1,3} and Collette Christophe\textsuperscript{1,3}  \thanks{\textsuperscript{1} Precision Mechatronics Laboratory, A\&M Department, University of Liege, Belgium} \thanks{\textsuperscript{2} European Synchrotron Radiation Facility, Grenoble, France} \thanks{\textsuperscript{3} BEAMS Department, Free University of Brussels, Belgium} \thanks{\textsuperscript{$\dagger$} Corresponding Author. Email: {\tt\small thomas.dehaeze@esrf.fr}}},
 pdftitle={Complementary Filters Shaping Using \(\mathcal{H}_\infty\) Synthesis},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.3 (Org mode 9.2.6)},
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org239b4b4}
\label{sec:introduction}
A set of filters is said to be complementary if the sum of their transfer functions is equal to one at all frequencies.
These filters are used when two or more sensors are measuring the same physical quantity with different noise characteristics. Unreliable frequencies of each sensor are filtered out by the complementary filters and then combined to form a super sensor giving a better estimate of the physical quantity over a wider bandwidth.
This technique is called sensor fusion and is used in many applications.\par
In \cite{zimmermann92_high_bandw_orien_measur_contr,corke04_inert_visual_sensin_system_small_auton_helic,min15_compl_filter_desig_angle_estim}, various sensors (accelerometers, gyroscopes, vision sensors, etc.) are merged using complementary filters for the attitude estimation of Unmanned Aerial Vehicles (UAV).
In \cite{collette15_sensor_fusion_method_high_perfor}, several sensor fusion configurations using different types of sensors are discussed in order to increase the control bandwidth of active vibration isolation systems.
Furthermore, sensor fusion is used in the isolation systems of the Laser Interferometer Gravitational-Wave Observator (LIGO) to merge inertial sensors with relative sensors
\cite{matichard15_seism_isolat_advan_ligo,hua04_polyp_fir_compl_filter_contr_system}. \par
As the super sensor noise characteristics largely depend on the complementary filter norms, their proper design is of primary importance for sensor fusion.
In \cite{corke04_inert_visual_sensin_system_small_auton_helic,jensen13_basic_uas,min15_compl_filter_desig_angle_estim}, first and second order analytical formulas of complementary filters have been presented.
Higher order complementary filters have been used in
\cite{shaw90_bandw_enhan_posit_measur_using_measur_accel,zimmermann92_high_bandw_orien_measur_contr,collette15_sensor_fusion_method_high_perfor}.
In \cite{jensen13_basic_uas}, the sensitivity and complementary sensitivity transfer functions of a feedback architecture have been proposed to be used as complementary filters. The design of such filters can then benefit from the classical control theory developments.
Linear Matrix Inequalities (LMIs) are used in \cite{pascoal99_navig_system_desig_using_time} for the synthesis of complementary filters satisfying some frequency-like performance.
Finally, a synthesis method of high order Finite Impulse Response (FIR) complementary filters using convex optimization has been developed in \cite{hua05_low_ligo,hua04_polyp_fir_compl_filter_contr_system}. \par
Although many design methods of complementary filters have been proposed in the literature, no simple method that allows to shape the norm of the complementary filters is available.
This paper presents a new design method of complementary filters based on \(\mathcal{H}_\infty\) synthesis.
This design method permits to easily shape the norms of the generated filters.\par
The section \ref{sec:requirements} gives a brief overview of sensor fusion using complementary filters and explains how the typical requirements for such fusion can be expressed as upper bounds on the filters norms.
In section \ref{sec:hinf_method}, a new design method for the shaping of complementary filters using \(\mathcal{H}_\infty\) synthesis is proposed.
In section \ref{sec:application_ligo}, the method is used to design complex complementary filters that are used for sensor fusion at the LIGO.
Our conclusions are drawn in the final section.

\section{Complementary Filters Requirements}
\label{sec:org35ebf13}
\label{sec:requirements}
\subsection{Sensor Fusion Architecture}
\label{sec:org291a5d3}
\label{sec:sensor_fusion}

Let's consider two sensors measuring the same physical quantity \(x\) with dynamics \(G_1(s)\) and \(G_2(s)\), and with uncorrelated noise characteristics \(n_1\) and \(n_2\).

@@

\subsection{Noise Sensor Filtering}
\label{sec:org9e972a6}
\label{sec:noise_filtering}

@@

Let's define the estimation error \(\delta x\) by \eqref{eq:estimate_error}.
\begin{equation}
\label{eq:estimate_error}
  \delta x \triangleq \hat{x} - x = H_1 n_1 + H_2 n_2
\end{equation}

@@

\subsection{Robustness of the Fusion}
\label{sec:org7c81675}
\label{sec:fusion_robustness}

In practical systems the sensor dynamics is not perfect and \eqref{eq:perfect_dynamics} is not verified.
In such case, one can use an inversion filter \(\hat{G}_i^{-1}(s)\) to normalize the sensor dynamics, where \(\hat{G}_i(s)\) is an estimate of the sensor dynamics \(G_i(s)\).
However, as there is always some level of uncertainty on the dynamics, it cannot be perfectly inverted and \(\hat{G}_i^{-1}(s) G_i(s) \neq 1\).

Let's represent the resulting dynamic uncertainty of the inverted sensors by an input multiplicative uncertainty as shown in Fig. \ref{fig:sensor_fusion_dynamic_uncertainty} where \(\Delta_i\) is any stable transfer function satisfying \(|\Delta_i(j\omega)| \le 1,\ \forall\omega\), and \(|w_i(s)|\) is a weight representing the magnitude of the uncertainty.

@@

The super sensor dynamics \eqref{eq:super_sensor_dyn_uncertainty} is no longer equal to \(1\) and now depends on the sensor dynamics uncertainty weights \(w_i(s)\) as well as on the complementary filters \(H_i(s)\).
\begin{equation}
\label{eq:super_sensor_dyn_uncertainty}
  \frac{\hat{x}}{x} = 1 + w_1(s) H_1(s) \Delta_1(s) + w_2(s) H_2(s) \Delta_2(s)
\end{equation}

The uncertainty region of the super sensor can be represented in the complex plane by a circle centered on \(1\) with a radius equal to \(|w_1(j\omega) H_1(j\omega)| + |w_2(j\omega) H_2(j\omega)|\) as shown in Fig. \ref{fig:uncertainty_set_super_sensor}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/uncertainty_set_super_sensor.pdf}
\caption{\label{fig:uncertainty_set_super_sensor}
Uncertainty region of the super sensor dynamics in the complex plane (solid circle). The contribution of both sensors 1 and 2 to the uncertainty are represented respectively by a dotted and a dashed circle}
\end{figure}

The maximum phase added \(\Delta\phi(\omega)\) by the super sensor dynamics at frequency \(\omega\) is then
\begin{equation}
\label{eq:max_phase_uncertainty}
    \Delta\phi(\omega) = \arcsin\big( |w_1(j\omega) H_1(j\omega)| + |w_2(j\omega) H_2(j\omega)| \big)
\end{equation}

As it is generally desired to limit the maximum phase added by the super sensor, \(H_1(s)\) and \(H_2(s)\) should be designed such that \eqref{eq:max_uncertainty_super_sensor} is satisfied.
\begin{equation}
\label{eq:max_uncertainty_super_sensor}
   \max_\omega \big( \left|w_1 H_1\right| + \left|w_2 H_2\right|\big) < \sin\left( \Delta \phi_\text{max} \right)
\end{equation}

Thus the norm of the complementary filter \(|H_i|\) should be made small at frequencies where \(|w_i|\) is large.

\section{Complementary Filters Shaping using \(\mathcal{H}_\infty\) Synthesis}
\label{sec:org4512d61}
\label{sec:hinf_method}
As shown in Sec. \ref{sec:requirements}, the performance and robustness of the sensor fusion architecture depends on the complementary filters norms.
Therefore, the development of a synthesis method of complementary filters that allows the shaping of their norm is necessary.
\subsection{Shaping of Complementary Filters using \(\mathcal{H}_\infty\) synthesis}
\label{sec:org42885ea}
\label{sec:hinf_synthesis}
The synthesis objective is to shape the norm of two filters \(H_1(s)\) and \(H_2(s)\) while ensuring their complementary property \eqref{eq:comp_filter}.
This is equivalent as to finding stable transfer functions \(H_1(s)\) and \(H_2(s)\) such that conditions \eqref{eq:comp_filter_problem_form} are satisfied.
\begin{subequations}
\label{eq:comp_filter_problem_form}
  \begin{align}
  & H_1(s) + H_2(s) = 1 \label{eq:hinf_cond_complementarity} \\
  & |H_1(j\omega)| \le \frac{1}{|W_1(j\omega)|} \quad \forall\omega \label{eq:hinf_cond_h1} \\
  & |H_2(j\omega)| \le \frac{1}{|W_2(j\omega)|} \quad \forall\omega \label{eq:hinf_cond_h2}
  \end{align}
\end{subequations}
where \(W_1(s)\) and \(W_2(s)\) are two weighting transfer functions that are chosen to shape the norms of the corresponding filters.

@@

\subsection{Weighting Functions Design}
\label{sec:org6492bf4}
\label{sec:hinf_weighting_func}
The proper design of the weighting functions is of primary importance for the success of the presented complementary filters \(\mathcal{H}_\infty\) synthesis.

@@

\subsection{Validation of the proposed synthesis method}
\label{sec:org1650523}
\label{sec:hinf_example}
Let's validate the proposed design method of complementary filters with a simple example where two complementary filters \(H_1(s)\) and \(H_2(s)\) have to be designed such that:
\begin{itemize}
\item the merging frequency is around \(\SI{10}{Hz}\)
\item the slope of \(|H_1(j\omega)|\) is \(-2\) above \(\SI{10}{Hz}\)
\item the slope of \(|H_2(j\omega)|\) is \(+3\) below \(\SI{10}{Hz}\)
\item the gain of both filters is equal to \(10^{-3}\) away from the merging frequency
\end{itemize}

@@

\subsection{Synthesis of Three Complementary Filters}
\label{sec:orga872169}
\label{sec:hinf_three_comp_filters}
Some applications may require to merge more than two sensors.
In such a case, it is necessary to design as many complementary filters as the number of sensors used.
The synthesis problem is then to compute \(n\) stable transfer functions \(H_i(s)\) such that \eqref{eq:hinf_problem_gen} is satisfied.
\begin{subequations}
\label{eq:hinf_problem_gen}
  \begin{align}
  & \sum_{i=0}^n H_i(s) = 1 \label{eq:hinf_cond_compl_gen} \\
  & \left| H_i(j\omega) \right| < \frac{1}{\left| W_i(j\omega) \right|}, \quad \forall \omega,\ i = 1 \dots n \label{eq:hinf_cond_perf_gen}
  \end{align}
\end{subequations}
The synthesis method is generalized here for the synthesis of three complementary filters using the architecture shown in Fig. \ref{fig:comp_filter_three_hinf}.

@@

\section{Application: Design of Complementary Filters used in the Active Vibration Isolation System at the LIGO}
\label{sec:org6b43181}
\label{sec:application_ligo}
Several complementary filters are used in the active isolation system at the LIGO \cite{hua05_low_ligo,hua04_polyp_fir_compl_filter_contr_system}.
The requirements on those filters are very tight and thus their design is complex.
The approach used in \cite{hua05_low_ligo} for their design is to write the synthesis of complementary FIR filters as a convex optimization problem.
The obtained FIR filters are compliant with the requirements. However they are of very high order so their implementation is quite complex.

The effectiveness of the proposed method is demonstrated by designing complementary filters with the same requirements as the one described in \cite{hua05_low_ligo}.
\subsection{Complementary Filters Specifications}
\label{sec:orge7d0e23}
\label{sec:ligo_specifications}
The specifications for one pair of complementary filters used at the LIGO are summarized below (for further details, refer to \cite{hua04_polyp_fir_compl_filter_contr_system}) and shown in Fig. \ref{fig:ligo_weights}:
\begin{itemize}
\item From \(0\) to \(\SI{0.008}{Hz}\), the magnitude of the filter's transfer function should be less or equal to \(8 \times 10^{-4}\)
\item Between \(\SI{0.008}{Hz}\) to \(\SI{0.04}{Hz}\), the filter should attenuate the input signal proportional to frequency cubed
\item Between \(\SI{0.04}{Hz}\) to \(\SI{0.1}{Hz}\), the magnitude of the transfer function should be less than \(3\)
\item Above \(\SI{0.1}{Hz}\), the magnitude of the complementary filter should be less than \(0.045\)
\end{itemize}

\subsection{Weighting Functions Design}
\label{sec:org03404b3}
\label{sec:ligo_weights}
The weighting functions should be designed such that their inverse magnitude is as close as possible to the specifications in order to not over-constrain the synthesis problem.
However, the order of each weight should stay reasonably small in order to reduce the computational costs of the optimization problem as well as for the physical implementation of the filters.

A Type I Chebyshev filter of order \(20\) is used as the weighting transfer function \(w_L(s)\) corresponding to the low pass filter.
For the one corresponding to the high pass filter \(w_H(s)\), a \(7^{\text{th}}\) order transfer function is designed.
The magnitudes of the weighting functions are shown in Fig. \ref{fig:ligo_weights}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/ligo_weights.pdf}
\caption{\label{fig:ligo_weights}
Specifications and weighting functions magnitudes}
\end{figure}

\subsection{\(\mathcal{H}_\infty\) Synthesis}
\label{sec:orgd133413}
\label{sec:ligo_results}
\(\mathcal{H}_\infty\) synthesis is performed using the architecture shown in Fig. \ref{eq:generalized_plant}.
The complementary filters obtained are of order \(27\).
In Fig. \ref{fig:comp_fir_ligo_hinf}, their bode plot is compared with the FIR filters of order 512 obtained in \cite{hua05_low_ligo}.
They are found to be very close to each other and this shows the effectiveness of the proposed synthesis method.

@@

\section{Conclusion}
\label{sec:orgde38c04}
  \label{sec:conclusion}
This paper has shown how complementary filters can be used to combine multiple sensors in order to obtain a super sensor.
Typical specification on the super sensor noise and on the robustness of the sensor fusion has been shown to be linked to the norm of the complementary filters.
Therefore, a synthesis method that permits the shaping of the complementary filters norms has been proposed and has been successfully applied for the design of complex filters.
Future work will aim at further developing this synthesis method for the robust and optimal synthesis of complementary filters used in sensor fusion.

\section{Acknowledgment}
\label{sec:org57822f5}
This research benefited from a FRIA grant from the French Community of Belgium.
